---
title: Exercises
format: 
  html:
    toc: false
date-modified: last-modified
date-format: long
summary: intro to workflows
---

```{r, echo = FALSE, results='asis'}
library(webexercises)

knitr::opts_chunk$set(echo = FALSE)

# Uncomment to change widget colours:
style_widgets(incorrect = "red", correct = "green", highlight = "firebrick")
```

Put your learning to the test with what you’ve covered so far.

**1. What role does a workflow manager play in computational research??**

```{r}
#| echo: false
opts_1 <- c(
   answer="Automating the execution of complex computational processes",
   "To manually execute each step of a computation",
   "Offering real-time and intuitive data analysis",
   "To minimize the reliance on specific software environments"
)
```

`r longmcq(opts_1)`

**2.What is the primary drawback of using shell scripts for automating computations?**

```{r}
#| echo: false
opts_2 <- c(
   answer="Limited error handling and debugging capabilities",
   "Difficulty in integrating with notebooks",
   "Not compatible with all operating systems",
   answer="They re-run all the steps every time",
   "Insufficient support for parallel processing",
   "Complex and extensive coding for simple tasks"
)
```

`r longmcq(opts_2)`


**3. What are the key features of workflow manager in computational research?**
(Several possible solutions)

```{r}
#| echo: false
opts_3 <- c(
   answer="Executing tasks only when required",
   answer = "Managing task dependencies",
   answer = "Overseeing storage and resource allocation",
   "Providing intuitive graphical interfaces"
)
```

`r longmcq(opts_3)`

**4. Workflow managers can run tasks (different) concurrently if there are no dependencies** (True or False) `r torf(TRUE)`

**5. A workflow manager can execute a single parallelized task on multiple nodes in a computing cluster** (True or False) `r torf(TRUE)`

